Dynamics of Anthropomorphism in Human-Robot Interaction

### 2014-04-07 paper.tex
refinements sections 1,2,3

\title[The Dynamics of Anthropomorphism]{The Dynamics of Anthropomorphism in Human-Robot Interaction}
\author[\firstAuthorLast ]{\Authors}
\address{}
\correspondance{}
\extraAuth{}
\topic{The Uncanny Valley Hypothesis and Beyond}% If your article is part of a Research Topic, please indicate here which.

@@

A phenomenon that is often studied related to the human-like design of 
robots is anthropomorphism. While people's tendency to perceive human-like 
characteristics in robots is a commonly discussed trait of 
human-robot interaction (HRI), the phenomenon paradoxically lacks formal 
grounds. Based on findings from several empirical studies and a 
comprehensive synthesis of literature, this article attempts at extending 
the current understanding of anthropomorphism. In accordance with theories 
from social psychology, cognitive science, and neuroscience, we suggest to 
formalize qualitative differences and dynamics over time in how 
humans anthropomorphize non-human agents. Based on several key factors, 
namely the 
design of the non-human agent, the individual characteristics of the human 
user, and the context of use, we found that anthropomorphism changes in a 
non-linear way with growing experience of the user with the system. To 
account for these variances, we propose an initial model that we call 
\textit{``Dynamics of Anthropomorphism''}. Focusing on human-robot 
interactions, we present a cognitive interpretation that considers 
the described dynamics. Finally, we would like to discuss how others can 
apply our model to measure and describe better the complexity in human's 
tendency to perceive human-like characteristics in non-human agents.

@@

The Uncanny Valley hypothesis \citep{mori_uncanny_1970} proposes that human-like
designs of objects and non-human agents can evoke both positive and negative feelings in the human observer / user. A positive valence reflects the experience of 
emotional engagement with and feelings of empathy for the non-human agent
\citep{cheetham_human_2011}. This in turn may lead to an effective (human-like and 
human-social) interaction with the non-human agent, and increase people's acceptance of
it. 
However, at a certain point as the non-human agent
approaches but fails to closely resemble a human, the evoked valence in the
human user suddenly turns into a negative uncanny feeling of what
\cite{mori_uncanny_1970} describes as disgust and revulsion. This, in turn,
might lead to rejection of the non-human agent. The Uncanny
Valley hypothesis predicts that with further increasing the human-likeness of
the non-human agent such that it appears so realistic that it is not
distinguishable from a human but perceived to be human, the effect is again a
positive one, reflecting feelings and engagement that are similar as those
evoked by a human-human interaction. 
Of course, developers and designers of artificial agents
(\eg virtual characters, robots) try to exploit the positive effects of 
\textbf{anthropomorphic form} and avoid ``falling'' into this Uncanny Valley.
Consequently, there is a considerable research effort to
better understand the boundaries of anthropomorphic design and their effects
on the interaction and experience. To date, findings have been however
inconsistent \citep{cheetham_human_2011}.

A phenomenon that is often studied related to the human-like design of objects and 
artificial agents is \textbf{anthropomorphism}.
Along this line, this article aims at fostering a better understanding of
anthropomorphism (noted \Ant in this paper) in human-robot interaction 
by considering the phenomenon
as a whole. Robotics researchers often tend to believe that
anthropomorphism only describes a static set of human-like features of a robot
(like shape, speech capabilities, facial expressions, etc.). We refer to these
characteristics as the \emph{anthropomorphic design} of the
robot~\citep{fink_anthropomorphism_2012}, and propose that \emph{anthropomorphism} in 
general refers to a \emph{social phenomenon} that emerges from the (real or
imagined) \emph{interaction} between a non-human agent (\eg a robot) and a 
human~\citep{persson_anthropomorphism_2000}. According to~\cite{epley_when_2008},
anthropomorphism includes that the human perceives, for instance, emotional states, motivations, or 
intentions in the non-human agent, and tends to ascribe those qualities to it.
As such, the dynamic and
socio-cognitive dimensions of anthropomorphism are essential to understand the
complex bonds that human users build with artificial agents such as robots.

@@

\begin{figure}[h!]
        \centering
        \begin{subfigure}[t]{0.48\columnwidth}
                \includegraphics[width=\textwidth]{roomba-activities.png}
                \caption{Activities related to Roomba, n = 9 households}
                \label{fig:roomba-activities}
        \end{subfigure}%
        \hspace{0.5cm} 
        \begin{subfigure}[t]{0.48\columnwidth}
                \includegraphics[width=\textwidth]{roomba-perception.pdf}
                \caption{People's perception of Roomba, n = 15 participants}
                \label{fig:roomba-perception}
        \end{subfigure}
    \caption{Usage and perception of the Roomba vacuum-cleaning robot over a 6 months period. Qualitative data suggest that people's activities related to the robot as well as their perception of the robot change over time.}
    \label{fig:roomba}
\end{figure}

@@

\begin{figure}[h!]
        \centering
        \begin{subfigure}[b]{0.48\columnwidth}
                \centering
                \includegraphics[width=0.75\textwidth]{ranger}
                \caption{Interaction scenario of the \emph{Ranger} domino study: the robot toy box Ranger transports domino tiles between the two children}
                \label{fig:ranger-expe}
        \end{subfigure}%
        \hspace{0.5cm}
        \begin{subfigure}[b]{0.48\columnwidth}
            \includegraphics[width=\textwidth]{ranger-interactions-vs-qualitative-score}
            \caption{We found a significant negative correlation ($p=0.048$) between the
    number of interactions and the qualitative anthropomorphic projections. n = 13 groups}
            \label{fig:qualitative-score}
        \end{subfigure}
        \caption{Children`s interaction with Ranger, and their anthropomorphic projections onto the robot. Overall, large variations of both children`s behavior and their perception of the robot were found but seemed to be correlated to each other.}
    \label{fig:ranger}
\end{figure}

@@

These first remarks lead to the main question we attempt to address in this
article: how can we understand \emph{anthropomorphism} as a whole, \hl{and in such a
way that is support improved human-robot interactions?}\fixme{JF: something's wrong with this sentence}

@@

@@

Literature on anthropomorphism is quite diverse. We would like to point out that there exist different and partly conflicting definitions of the term anthropomorphism, and that in general, the \textit{terminology} is not clear. Probably as a result of this, there also exist different understandings of the \textit{phenomenon} that the term `anthropomorphism' describes.
 In this section we will first give an overview on the diversity of terms and
 then try to establish a common ground in the understanding of anthropomorphism
 as a psychological phenomenon.

@@

Originally, the term \textbf{anthropomorphism} comes from the Greek
\textit{anthropos} for ``man'' (or ``human'') and \textit{morphe} for `form /
structure' (or `shape'). Anthropomorphism is generally understood as people's
tendency to perceive a non-human agent as if it had some human characteristics
or traits. How a person perceives an agent is crucial, as it further determines
how he/she relates to, and behaves towards the agent, or how he/she may behave in
light of this agent \citep{epley_when_2008}. An anthropomorphic relationship is
often characterized by the fact that the human attributes human-unique
characteristics to the non-human agent. Since anthropomorphism becomes
observable through these attributions, some definitions claim that
anthropomorphism is not the \textit{perception} but rather the
\textit{attribution} of human qualities to non-human agents.  Ascribing
thoughts, emotions, and intentions are given as examples, and as the most
prominently anthropomorphized non-human agents, animals and robots are mentioned
\citep{duffy_anthropomorphism_2003,schmitz_concepts_2011}.

Apart from anthropomorphism, there also exists the term \textbf{animism}, which
can be defined as the attribution of life to the nonliving. The study of animism
was pioneered by Piaget in the 1920s with his work on \textit{The Child's
Conception of the  World}. He found that children tend to regard many inanimate
objects as capable of sensations, emotions and intentions, and called this view
``animism''. Since then, children's inferences about the world have been studied
intensively, and there is an agreement that children's views and concepts of
living, animal, and plant evolve as they grow up. Often synonymously used to
animism is the term \textbf{zoomorphism}, which similarly applies to the cases
when non-lifelike instances / artifacts are associated with animalistic
attributes, excluding human-specific traits.  As ``being alive'' is not a
human-unique property but applies to all living beings, the attribution of life
may not be labeled as anthropomorphism (in a narrow understanding of it).
Another possibility to illustrate how anthropomorphism differs from attributions
of animacy is to look at paraphrases of anthropomorphism. That is to say,
anthropomorphism can alternatively be understood as \textbf{humanization} or
\textbf{personification} of anything other than a human being\footnote{There
exist also the phenomenons inverse to anthropomorphism, which can be found
in the literature as \textbf{dehumanization}
\citep{haslam_dehumanization:_2006}, or \textbf{mechanomorphism}
\citep{caporael_anthropomorphism_1986}.}. Robots can
both be animized and anthropomorphized, but often the difference is not made.

@@

Several attempts have been made to clarify what the psychological phenomenon
\emph{anthropomorphism} actually is.  From a psychology point of view,
\citet{epley_when_2008} explain what anthropomorphism is \textit{not}. For
instance, in their understanding, anthropomorphism does not include behavioral
descriptions of observable actions but it requires going beyond what is directly
observable. They also suggest that anthropomorphism does not merely entail
\textit{animism}, as animate life is not a \textit{uniquely} human-like
characteristic\footnote{For the whole list and more details, the reader may
refer to the original work \citep{epley_when_2008}.}.

@@

In the following we will only briefly present the artifact-centered perspective to explain anthropomorphism\footnote{A review of anthropomorphic design of robots can also be found in \cite{fink_anthropomorphism_2012}}. Then, we provide a more detailed review and synthesis of literature that applies a human-centered perspective to explain anthropomorphism. Concretely, we integrate theories from developmental psychology, cognitive science / neuroscience, and social psychology to shed light on the question why humans anthropomorphize.

@@

The direction of causality in the artifact-centered view goes from the artifact
to the humans: the artifact (or a certain feature of it) \emph{encourages} the
human to anthropomorphize it.  This perspective assumes that humans directly
(mindlessly) respond to life-like and social cues that the non-human agent or
system emits \citep{nass_machines_2000}. Without thoughtful mental processing,
humans tend to simply apply stereotypes and heuristics to the non-human agent,
and in turn apply human-human social schemas and norms to the occurring
interactions.  \cite{takayama_perspectives_2012} propose to refer to this
immediate (sometimes visceral) sense in a situation as \textit{in-the-moment},
compared to a more distanced cogitation and consideration which she calls
\textit{reflective}. As this distinction implies different cognitive processes,
it is more approriate to apply a human-centered perspective and we will come
back to it later, in section \ref{sec:cognitive-expl}.

@@

Humans tend to perceive human-like characteristics such as physical appearance,
emotional states, or inner mental states and motivations in non-human agents
that do not have these characteristics themselves, such as animals, natural
forces, religious agents, technological gadgets, or mechanical devices
\citep{epley_when_2008}\footnote{According to \citet{epley_when_2008}, a
non-human agent can be anything that acts -- or is believed to act -- with
apparent independence.}. One common reason for anthropomorphism is that by
perceiving some human-like characteristics in the non-human agent, this agent is
made more \emph{graspable}, \emph{understandable}, \emph{predictable}, and one
can feel \emph{more empathetic} toward it, and possibly one can then anticipate
the agent's behavior. These are central motivations when trying to explain
anthropomorphism, and we will come back to them in more detail later on.
\citet{eddy_attribution_1993} explain that, because of these reasons, people
commonly anthropomorphize their pets, by ascribing mental states, intentions or
feelings to them\footnote{Interestingly, pet-ownership seems to have a
significant impact on human-robot interaction: several studies suggest that
pet-owners are more likely to anthropomorphize technologies and robots than
non pet-owners}. But how can these motivations behind anthropomorphism be
explained from a human-centered point of view?

@@

The phenomenon of attributing animacy and intentions to nonliving objects such
as simple shapes has been intensively studied in developmental psychology. As
mentioned earlier, Piaget found that children tend to ascribe life to the
nonliving. They also tend to interpret physical phenomena in terms of intention
on the part of the non-human subject. For instance children are likely to argue
that \textit{``the sun is hot because it wants to make people warm''}
\citep{leeds_childrens_1992}. In this example it is in fact questionable whether
children really anthropomorphize the sun (believing that it has an intention) or
whether they are using anthropomorphism in a metaphoric sense (because their
conceptions of the world and of the living are not yet fully formed). Later
however, two famous experiments by \cite{heider_experimental_1944} and
\cite{michotte_perception_1963} showed that also adult participants attribute
animacy and intention to simple things and shapes based on motion. In both
experiments, participants viewed animations of simple shapes, such as circles or
triangles. Asked to describe what they observed, most people developed
elaborate stories, attributing motivations, emotions and relationships between
the objects\footnote{For more details, the reader may refer to the ``attribution
theory''.}.

@@

This cognitive viewpoint on anthropomorphism is important because people's
estimation of an agent's knowledge model and its capabilities affects the way
they relate to it. This holds implications for the resulting interaction with
the system, the user experience, and the acceptance. Previous research examined
the validity of the mental model concept with various kinds of robots
\citep{schmitz_concepts_2011,kiesler_mental_2002}. Findings suggest that people
tend to hold richer mental models of anthropomorphic robots in contrast to
mechanic ones \citep{kiesler_mental_2002}\footnote{Later in this article,
we may however come to question how far this result is based on
expectations from the human user side, which are likely to be refined
after continued interaction, and after the user gets acquainted with the
robot's behavior.}. A similar finding is described in
\cite{hegel_understanding_2008} and \cite{krach_can_2008}. In a user study using
functional magnetic resonance imagery (fMRI), the authors found that
participants implicitly attributed human-like qualities, such as mental states,
to their non-human interaction partner. Indeed, from an early age humans develop
a tendency to explain one's own and others' actions in terms of beliefs, desires
and goals,  called \textit{theory of mind}. The theory of mind allows the
implicit attribution of intentions and other mental states to others
\citep{premack1978does,leslie_pretense_1987,Frith2003}.  The finding of
\cite{hegel_understanding_2008} and \cite{krach_can_2008} was evident at the
behavioral level as well as on the neuro-physiological level: the more the
human-likeness of the artefactual partner increases, the more brain areas
associated with theory of mind  get activated \citep{krach_can_2008}.
Implication of human brain area involved in the inference of others` mental
states has also been shown in response to viewing
non-anthropomorphic/non-humanoid robotic gadgets whose behavior has been
described as unpredictable -- but not in response to those whose behavior was
described as predictable -- a finding also found at the behavioral level
\citep{Waytz2010}.  Interestingly, the grey matter volume of a brain area
related to theory of mind has been correlated to individual's score of
anthropomorphism \citep{cullen2013individual}.

At lower, more automatic, level, robots have been shown to elicit resonance
behaviors in the human brain. Resonance behaviors \citep{Rizzolatti1999} are
mechanisms by which the brain areas involved in the internal representations of
an action, an emotion or a sensation are equally recruited during the perception
of another individual performing the same action or experiencing the same
emotion or sensation.  The neurons showing resonant properties have been called
mirror neurons and have been found in a wide range of modalities (motor,
emotional etc.) \fixme{addref}. Seeing a robotic arm reaching for an object
activates the motor mirror neurons the same way as for seeing a human arm
reaching for the same object \citep{Gazzola2007, oberman_eeg_2007}. However, a
study of emotion perception on a robotic face found reduced activity in
emotional brain area known to have mirror properties in response to the robotic
face compared to a human face \citep{Chaminade2010}. Nevertheless, these authors
showed that when the participants were explicitly instructed to pay attention to
the robot's emotional expression, motor mirror neurons get activated.
Interestingly, using electroencephalography,  it has also been shown that
emotional behavior elicited by a robot face influences the speed of human
responses as well as early brain processes the same way as emotional expression
of a human face did \citep{Dubal2010}.

In sum, this branch of research suggests that anthropomorphism implies that
people thoughtfully develop a mental model of agents in their environment and
that they make inferences about these agents based on what is familiar to them
-- themselves, humans and human behavior. This understanding builds on the
theory of mind,  a
person's ability to attribute mental states to oneself and others. 
%A theory of mind for other agents enables us to attribute intentionality to those agents \citep{leslie_pretense_1987,admoni_multi-category_2012}. 
We briefly reviewed  neuro-physiological evidence of a link between the tendency
to thoughtfully anthropomorphize and the engagement of brain processes involved
in the attribution of mental states to other humans. The human-likeness quality
of the artefactual agent seems to play a role in the involvement of theory of
mind but also the unpredictable feature of the agent behavior alone seems to be
sufficient. Low-level mechanisms of the brain that allow humans to map others`
motor behaviors into their own repertoire are equally elicited when perceiving a
robot's action. It is not yet clear whether humans are processing emotional
signals from a robot the same way as they do for human beings. 
%were recently demonstrated at the brain level~\citep{cullen2013individual}.

@@

The first factor (\textit{elicited agent knowledge}) is a \emph{cognitive
determinant} of anthropomorphism and based on the idea that a person builds a
mental model (theory of mind) of the other agent. \citet{epley_seeing_2007}
suggest that the process of making inferences about non-human agents is based on
the activation of knowledge about humans (or the self). As mentioned before,
when a person builds a mental model of the other (unfamiliar) agent, the
question here is why does she draw on knowledge about other humans / herself and
not on something else? One basic reasons for this is a person's physical
constraints of being a human and nothing else. Consequently, one has no other
experience than the self and in turn people tend to make inferences about
others' mental states by relying inordinately on their own mental state
\footnote{Using one's own mental states and characteristics as a guide when
reasoning about other humans is called ego-centrism. In contrast, when using
self-knowledge (or knowledge about humans in general) when reasoning about
non-human agents, this is called anthropomorphism. See
\cite{epley_seeing_2007}}.  Also empirical findings suggest that knowledge about
humans in general, or self-knowledge in particular, is likely to serve as a
readily accessible base for induction when reasoning about non-human agents.
This tendency is usually stronger in young children and decreases with cognitive
development and the learning to distinguish the self from other humans, and
non-human agents.

@@

To facilitate things, we introduce the notation \Ant[R,H,C,t] to bring these multiple factors together. To take into account dynamics over time, we suggest that
anthropomorphism is a function of a vector of robot's traits $R = \{r_0,...,r_n\}$, a vector human
determinants $H = \{h_0,...,h_n\}$, a set of interaction situations that build
an interaction context $C = \{s_0,...,s_n\}$, and the duration / history of the
interaction itself, $t$.

>>>>

\title[Dynamics of Anthropomorphism]{Dynamics of Anthropomorphism in Human-Robot Interaction}
\author[\firstAuthorLast ]{\Authors}
\address{}
\correspondance{}
\extraAuth{}
\topic{The Uncanny Valley Hypothesis and Beyond}% If your article is part of a Research Topic, please indicate here which.

@@

A phenomenon that is often studied related to the human-like design of 
robots is \textit{anthropomorphism} -- people's tendency to perceive human-like 
characteristics in non-human agents. While this is a commonly discussed trait of 
human-robot interaction (HRI), the phenomenon paradoxically appears to lack formal 
grounds. Based on findings from several empirical studies and a 
comprehensive synthesis of literature, this article attempts at extending 
the current understanding of anthropomorphism. In accordance with theories 
from social psychology, cognitive science, and neuroscience, we suggest to 
formalize qualitative differences and dynamics over time in how 
humans anthropomorphize non-human agents. We propose several key factors that determine anthropomorphism, 
namely the 
design of the non-human agent, the individual characteristics of the human 
user, and the context of use. An initial model that we call 
\textit{``Dynamics of Anthropomorphism''} tries to formalize the qualitative changes in anthropomorphism over time. To 
account for these variances, our model describes the non-linear relationship between anthropomorphic effects of a human-robot interaction and the human user's growing experience with the system.
We also present a cognitive interpretation that considers 
these described dynamics. Further, we discuss scenarios and techniques to measure anthropomorphism. Finally, we make suggestions how others can 
apply our model to better describe the complexity in human's 
tendency to perceive human-like characteristics in non-human agents.

@@

The Uncanny Valley hypothesis \citep{mori_uncanny_1970} proposes that human-like
designs of objects and non-human agents can evoke both positive and negative feelings in the human observer / user. A positive valence reflects the experience of 
emotional engagement with and feelings of empathy for the non-human agent
\citep{cheetham_human_2011}. This in turn may lead to an effective (human-like and 
human-social) interaction with the non-human agent, and increase people's acceptance of
it. 
%
%However, at a certain point as the non-human agent
%approaches but fails to closely resemble a human, the evoked valence in the
%human user suddenly turns into a negative uncanny feeling of what
%\cite{mori_uncanny_1970} describes as disgust and revulsion. This, in turn,
%might lead to rejection of the non-human agent. The Uncanny
%Valley hypothesis predicts that with further increasing the human-likeness of
%the non-human agent such that it appears so realistic that it is not
%distinguishable from a human but perceived to be human, the effect is again a
%positive one, reflecting feelings and engagement that are similar as those
%evoked by a human-human interaction. 
%Of course, developers and designers of artificial agents
%(\eg virtual characters, robots) try to exploit the positive effects of 
%\textbf{anthropomorphic form} and avoid ``falling'' into this Uncanny Valley.
%Consequently, there is a considerable research effort to
%better understand the boundaries of anthropomorphic design and their effects
%on the interaction and experience. To date, findings have been however
%inconsistent \citep{cheetham_human_2011}.

A phenomenon that is often studied related to this is \textbf{anthropomorphism}.
Along this line, this article aims at fostering a better understanding of
anthropomorphism (noted \Ant in this paper) in human-robot interaction 
by considering the phenomenon
as a whole. Robotics researchers often tend to believe that
anthropomorphism only describes a static set of human-like features of a robot
(like shape, speech capabilities, facial expressions, etc.). We refer to these
characteristics as the \emph{anthropomorphic design} of the
robot~\citep{fink_anthropomorphism_2012}, and propose that \emph{anthropomorphism} in 
general refers to a \emph{social phenomenon} that emerges from the (real or
imagined) \emph{interaction} between a non-human agent (\eg a robot) and a 
human~\citep{persson_anthropomorphism_2000}. According to~\cite{epley_when_2008},
anthropomorphism includes that the human perceives, for instance, emotional states, motivations, or 
intentions in the non-human agent, and tends to ascribe those qualities to it.
As such, the dynamic and
socio-cognitive dimensions of anthropomorphism are essential to understand the
complex bonds that human users build with artificial agents such as robots.

@@

\begin{figure}[b]
        \centering
        \begin{subfigure}[t]{0.48\columnwidth}
                \includegraphics[width=\textwidth]{roomba-activities.png}
                \caption{Activities related to Roomba, n = 9 households}
                \label{fig:roomba-activities}
        \end{subfigure}%
        \hspace{0.5cm} 
        \begin{subfigure}[t]{0.48\columnwidth}
                \includegraphics[width=\textwidth]{roomba-perception.pdf}
                \caption{People's perception of Roomba, n = 15 participants}
                \label{fig:roomba-perception}
        \end{subfigure}
    \caption{Usage and perception of the Roomba vacuum-cleaning robot over a 6 months period. Qualitative data suggest that people's activities related to the robot as well as their perception of the robot change over time.}
    \label{fig:roomba}
\end{figure}

@@

\begin{figure}[b]
        \centering
        \begin{subfigure}[b]{0.48\columnwidth}
                \centering
                \includegraphics[width=0.75\textwidth]{ranger}
                \caption{Interaction scenario of the \emph{Ranger} domino study: the robot toy box Ranger transports domino tiles between the two children}
                \label{fig:ranger-expe}
        \end{subfigure}%
        \hspace{0.5cm}
        \begin{subfigure}[b]{0.48\columnwidth}
            \includegraphics[width=\textwidth]{ranger-interactions-vs-qualitative-score}
            \caption{We found a significant negative correlation ($p=0.048$) between the
    number of interactions and the qualitative anthropomorphic projections. n = 13 groups}
            \label{fig:qualitative-score}
        \end{subfigure}
        \caption{Children's interaction with Ranger, and their anthropomorphic projections onto the robot. Overall, large variations of both children's behavior and their perception of the robot were found but seemed to be correlated to each other.}
    \label{fig:ranger}
\end{figure}

@@

These first remarks lead to the main question we attempt to address in this
article: how can we understand \emph{anthropomorphism} as a whole, and in such a
way that it supports or improves human-robot interaction?

@@

Further, we would like to investigate the link between the variances in anthropomorphism and the correlating cognitive states.

@@

Literature on anthropomorphism is quite diverse. There exist different and partly conflicting definitions of the term anthropomorphism, and that in general, the \textit{terminology} is not clear. Probably as a result of this, there also exist different understandings of the \textit{phenomenon} that the term ``anthropomorphism'' describes.
 In this section we will first give an overview on the diversity of terms and
 then try to establish a common ground in the understanding of anthropomorphism
 as a psychological phenomenon.

@@

Originally, the term \textbf{anthropomorphism} comes from the Greek
\textit{anthropos} for ``man'' (or ``human'') and \textit{morphe} for ``form /
structure'' (or ``shape''). Anthropomorphism is generally understood as people's
tendency to perceive a non-human agent as if it had some human characteristics
or traits. How a person perceives an agent is crucial, as it further determines
how he/she relates to, and behaves towards the agent, or how he/she may behave in
light of this agent \citep{epley_when_2008}. An anthropomorphic relationship is
often characterized by the fact that the human attributes human-unique
characteristics to the non-human agent. Since anthropomorphism becomes
observable through these attributions, some definitions claim that
anthropomorphism is not the \textit{perception} but rather the
\textit{attribution} of human qualities to non-human agents.  Ascribing
thoughts, emotions, and intentions are given as examples, and as the most
prominently anthropomorphized non-human agents, animals and robots are mentioned
\citep{duffy_anthropomorphism_2003,schmitz_concepts_2011}.

Apart from anthropomorphism, there also exists the term \textbf{animism}, which
can be defined as the attribution of life to the nonliving. The study of animism
was pioneered by Piaget in the 1920s with his work on \textit{The Child's
Conception of the  World}. He found that children tend to regard many inanimate
objects as capable of sensations, emotions and intentions, and called this view
``animism''. Since then, children's inferences about the world have been studied
intensively, and there is an agreement that children's views and concepts of
living, animal, and plant evolve as they grow up. Often synonymously used to
animism is the term \textbf{zoomorphism}, which similarly applies to the cases
when non-lifelike instances / artifacts are associated with animalistic
attributes, excluding human-specific traits.  As ``being alive'' is not a
human-unique property but applies to all living beings, the attribution of life
may not be labeled as anthropomorphism (in a narrow understanding of it).
Another possibility to illustrate how anthropomorphism differs from attributions
of animacy is to look at paraphrases of anthropomorphism. That is to say,
anthropomorphism can alternatively be understood as \textbf{humanization} or
\textbf{personification} of anything other than a human being.\footnote{There
exist also the phenomenons inverse to anthropomorphism, which can be found
in the literature as \textbf{dehumanization}
\citep{haslam_dehumanization:_2006}, or \textbf{mechanomorphism}
\citep{caporael_anthropomorphism_1986}.} Robots can
both be animized and anthropomorphized, but often the difference is not made.

@@

Several attempts have been made to clarify what the psychological phenomenon
\emph{anthropomorphism} actually is.  From a psychology point of view,
\citet{epley_when_2008} explain what anthropomorphism is \textit{not}. For
instance, in their understanding, anthropomorphism does not include behavioral
descriptions of observable actions but it requires going beyond what is directly
observable. They also suggest that anthropomorphism does not merely entail
\textit{animism}, as animate life is not a \textit{uniquely} human-like
characteristic.\footnote{For the whole list and more details, the reader may
refer to the original work \citep{epley_when_2008}.}

@@

In the following we will only briefly present the artifact-centered perspective to explain anthropomorphism. A review of anthropomorphic design of robots can also be found in \cite{fink_anthropomorphism_2012}. Then, we provide a more detailed review and synthesis of literature that applies a human-centered perspective to explain anthropomorphism. Concretely, we integrate theories from developmental psychology, cognitive science / neuroscience, and social psychology to shed light on the question why humans anthropomorphize.

@@

The direction of causality in the artifact-centered view goes from the artifact
to the humans: the artifact (or a certain feature of it) \emph{encourages} the
human to anthropomorphize it.  This perspective assumes that humans directly
(mindlessly) respond to life-like and social cues that the non-human agent or
system emits \citep{nass_machines_2000}. Without thoughtful mental processing,
humans tend to simply apply stereotypes and heuristics to the non-human agent,
and in turn apply human-human social schemas and norms to the occurring
interactions.  \cite{takayama_perspectives_2012} proposes to refer to this
immediate (sometimes visceral) sense in a situation as \textit{in-the-moment},
compared to a more distanced cogitation and consideration which she calls
\textit{reflective}. As this distinction implies different cognitive processes,
it is more approriate to apply a human-centered perspective and we will come
back to it later, in section \ref{sec:cognitive-expl}.

@@

Humans tend to perceive human-like characteristics such as physical appearance,
emotional states, or inner mental states and motivations in non-human agents
that do not have these characteristics themselves, such as animals, natural
forces, religious agents, technological gadgets, or mechanical devices
\citep{epley_when_2008}.\footnote{According to \citet{epley_when_2008}, a
non-human agent can be anything that acts -- or is believed to act -- with
apparent independence.} One common reason for anthropomorphism is that by
perceiving some human-like characteristics in the non-human agent, this agent is
made more \emph{graspable}, \emph{understandable}, \emph{predictable}, and one
can feel \emph{more empathetic} toward it, and possibly one can then anticipate
the agent's behavior. These are central motivations when trying to explain
anthropomorphism, and we will come back to them in more detail later on.
\citet{eddy_attribution_1993} explain that, because of these reasons, people
commonly anthropomorphize their pets, by ascribing mental states, intentions or
feelings to them.\footnote{Interestingly, pet-ownership seems to have a
significant impact on human-robot interaction: several studies suggest that
pet-owners are more likely to anthropomorphize technologies and robots than
non pet-owners} But how can these motivations behind anthropomorphism be
explained from a human-centered point of view?

@@

The phenomenon of attributing animacy and intentions to nonliving objects such
as simple shapes has been intensively studied in developmental psychology. As
mentioned earlier, Piaget found that children tend to ascribe life to the
nonliving. They also tend to interpret physical phenomena in terms of intention
on the part of the non-human subject. For instance children are likely to argue
that \textit{``the sun is hot because it wants to make people warm''}
\citep{leeds_childrens_1992}. In this example it is in fact questionable whether
children really anthropomorphize the sun (believing that it has an intention) or
whether they are using anthropomorphism in a metaphoric sense (because their
conceptions of the world and of the living are not yet fully formed). However, experiments by \cite{heider_experimental_1944} and
\cite{michotte_perception_1963} showed that also adult participants attribute
animacy and intention to simple things and shapes based on motion. In both
experiments, participants viewed animations of simple shapes, such as circles or
triangles. Asked to describe what they observed, most people developed
elaborate stories, attributing motivations, emotions and relationships between
the objects.\footnote{For more details, the reader may refer to the ``attribution
theory''.}

@@

This cognitive viewpoint on anthropomorphism is important because people's
estimation of an agent's knowledge model and its capabilities affects the way
they relate to it. This holds implications for the resulting interaction with
the system, the user experience, and the acceptance. Previous research examined
the validity of the mental model concept with various kinds of robots
\citep{schmitz_concepts_2011,kiesler_mental_2002}. Findings suggest that people
tend to hold richer mental models of anthropomorphic robots in contrast to
mechanic ones \citep{kiesler_mental_2002}.\footnote{Later in this article,
we may however come to question how far this result is based on
expectations from the human user side, which are likely to be refined
after continued interaction, and after the user gets acquainted with the
robot's behavior.} A similar finding is described in
\cite{hegel_understanding_2008} and \cite{krach_can_2008}. In a user study using
functional magnetic resonance imagery (fMRI), the authors found that
participants implicitly attributed human-like qualities, such as mental states,
to their non-human interaction partner. Indeed, from an early age humans develop
a tendency to explain one's own and others' actions in terms of beliefs, desires
and goals,  called \textit{theory of mind}. The theory of mind allows the
implicit attribution of intentions and other mental states to others
\citep{premack1978does,leslie_pretense_1987,Frith2003}.  The finding of
\cite{hegel_understanding_2008} and \cite{krach_can_2008} was evident at the
behavioral level as well as on the neuro-physiological level: the more the
human-likeness of the artefactual partner increases, the more brain areas
associated with theory of mind  get activated \citep{krach_can_2008}.
Implication of human brain area involved in the inference of others' mental
states has also been shown in response to viewing
non-anthropomorphic/non-humanoid robotic gadgets whose behavior has been
described as unpredictable -- but not in response to those whose behavior was
described as predictable -- a finding also found at the behavioral level
\citep{Waytz2010}.  Interestingly, the grey matter volume of a brain area
related to theory of mind has been correlated to individual's score of
anthropomorphism \citep{cullen2013individual}.

At lower, more automatic, level, robots have been shown to elicit resonance
behaviors in the human brain. Resonance behaviors \citep{Rizzolatti1999} are
mechanisms by which the brain areas involved in the internal representations of
an action, an emotion or a sensation are equally recruited during the perception
of another individual performing the same action or experiencing the same
emotion or sensation.  The neurons showing resonant properties have been called
mirror neurons and have been found in a wide range of modalities (motor,
emotional \etc). Seeing a robotic arm reaching for an object
activates the motor mirror neurons the same way as for seeing a human arm
reaching for the same object \citep{Gazzola2007, oberman_eeg_2007}. However, a
study of emotion perception on a robotic face found reduced activity in
emotional brain area known to have mirror properties in response to the robotic
face compared to a human face \citep{Chaminade2010}. Nevertheless, these authors
showed that when the participants were explicitly instructed to pay attention to
the robot's emotional expression, motor mirror neurons get activated.
Interestingly, using electroencephalography,  it has also been shown that
emotional behavior elicited by a robot face influences the speed of human
responses as well as early brain processes the same way as emotional expression
of a human face did \citep{Dubal2010}.

In sum, this branch of research suggests that anthropomorphism implies that
people thoughtfully develop a mental model of agents in their environment and
that they make inferences about these agents based on what is familiar to them
-- themselves, humans and human behavior. This understanding builds on the
theory of mind,  a
person's ability to attribute mental states to oneself and others. 
%A theory of mind for other agents enables us to attribute intentionality to those agents \citep{leslie_pretense_1987,admoni_multi-category_2012}. 
We briefly reviewed  neuro-physiological evidence of a link between the tendency
to thoughtfully anthropomorphize and the engagement of brain processes involved
in the attribution of mental states to other humans. The human-likeness quality
of the artefactual agent seems to play a role in the involvement of theory of
mind but also the unpredictable feature of the agent behavior alone seems to be
sufficient. Low-level mechanisms of the brain that allow humans to map others'
motor behaviors into their own repertoire are equally elicited when perceiving a
robot's action. It is not yet clear whether humans are processing emotional
signals from a robot the same way as they do for human beings. 
%were recently demonstrated at the brain level~\citep{cullen2013individual}.

@@

The first factor (\textit{elicited agent knowledge}) is a \emph{cognitive
determinant} of anthropomorphism and based on the idea that a person builds a
mental model (theory of mind) of the other agent. \citet{epley_seeing_2007}
suggest that the process of making inferences about non-human agents is based on
the activation of knowledge about humans (or the self). As mentioned before,
when a person builds a mental model of the other (unfamiliar) agent, the
question here is why does she draw on knowledge about other humans / herself and
not on something else? One basic reasons for this is a person's physical
constraints of being a human and nothing else. Consequently, one has no other
experience than the self and in turn people tend to make inferences about
others' mental states by relying inordinately on their own mental state.
\footnote{Using one's own mental states and characteristics as a guide when
reasoning about other humans is called ego-centrism. In contrast, when using
self-knowledge (or knowledge about humans in general) when reasoning about
non-human agents, this is called anthropomorphism. See
\cite{epley_seeing_2007}.} Also empirical findings suggest that knowledge about
humans in general, or self-knowledge in particular, is likely to serve as a
readily accessible base for induction when reasoning about non-human agents.
This tendency is usually stronger in young children and decreases with cognitive
development and the learning to distinguish the self from other humans, and
non-human agents.

@@

To facilitate things, we introduce the notation \Ant[R,H,C,t] to bring these multiple factors together. To take into account dynamics over time, we suggest that
anthropomorphism is a function of a vector of robot's traits $R = \{r_0,...,r_n\}$, a vector human
determinants $H = \{h_0,...,h_n\}$, characteristics of the interaction context $C = \{c_0,...,c_n\}$, and the duration / history of the
interaction itself, $t$.
